Toxic Code Snippets on Stack Overflow

### 2018-05-18 cloverflow-tse/cloverflow.tex
Continued revising the TSE paper.

Toxic code snippets mean code snippets that are harmful for reuse and, in
several cases, are caused by online code cloning. Stack Overflow code snippets
cloned from open source software or online sources can become toxic when they
are (1)~outdated or (2)~violating their original software
license. Outdated code snippets
can be harmful since they are not up-to-date with their originals and may
contain defects. Code snippets from open source projects usually fall under a
specific software license, e.g.\ GNU General Public License (GPL). If they are
cloned to Stack Overflow answers without the license, and then flow to other projects
with conflicting licenses, legal issues may occur.

@@

\begin{table}
	\centering
	\caption{Stack Overflow and Qualitas datasets}
	\label{tab:datasets}
	%\small
	\begin{tabular}{lrrr}
		\toprule
		Data set & No. of files & SLOC & Median \\
		\midrule
		Stack Overflow & 72,365 & 1,840,581 & 17 \\ 
		Qualitas &  166,709 & 19,614,083 & 60,667 \\ 
		\bottomrule
	\end{tabular} 
\end{table}

@@

To search for outdated online code clones, we focused on the \textbf{QS} clone
pairs that were cloned from Qualitas to Stack Overflow and compared them with
their latest versions. We downloaded the latest version of the Qualitas projects
from their repositories on October 3, 2017. For each \textbf{QS} online clone
pair, we used the clone from Qualitas as a proxy. We searched for its latest
version by the file name and located the cloned region in the file based on the
method name or specific code statements. We then compared the Stack Overflow
snippet to its latest version line-by-line to find if any change has been made
to the source code. We also made sure that the changes did not come from the
modifications made to the Stack Overflow snippets by the posters but from the
updates in the projects themselves. When we found inconsistent lines between the
two versions, we used {\small\texttt{git blame}} to see who modified those lines
of code and the timestamps. We also collected IDs of issue tracking information
if the code change is linked to an automatic issue tracking system, such as Jira
or BugZilla.

@@

The visitors survey confirms the findings from the previous studies that Stack
Overflow code snippets can be problematic~\cite{Zhang2018,Acar2016,An2017,Fischer2017}. Sixty-six
percent of the visitors experienced a problem from reusing Stack Overflow code
snippets ranging from incorrect solutions, outdated solutions, mismatched
solutions, to buggy code. Although they are aware of the problems, half of them
(56\%) never reports back to the Stack Overflow discussions. On the other hand,
the visitors rarely give attributions to Stack Overflow when they reuse code
snippets from the website, similar to the findings reported
by~\cite{Baltes2017}. The visitors are generally not aware of the CC BY-SA 3.0
license, and more than half of them never check for license compatibility when
reusing Stack Overflow code snippets. 
Sixty-nine Stack Overflow visitors (79\%) who adopted code from Stack Overflow never 
check if the code snippet originated from a different source (e.g. an open source project) 
with an incompatible license to their projects.
Moreover, we also found that 9\% of the participants
encountered legal issues by copying code from Stack Overflow. To the best of our
knowledge, a study of legal problems from reusing Stack Overflow code snippets
has never been done before and will be our future work.

@@

Moreover, using the information in git commit messages, we can associate each
change to its respective issues in an issue tracking system, such as Bugzilla or
Jira. We found that in 58 cases, the cloned code snippets on Stack Overflow were
changed because of a request in the issue tracking system. Since issue tracking
systems are also used, besides bug reports, for feature request and feature
enhancements, having an issue tracking ID can reflect that the change is important
and not only a superficial fix such as code formatting.

\Cref{tab:stale_code_details} shows examples of the outdated online clones on
Stack Overflow. The table displays information of the clones from both Stack
Overflow and Qualitas side including the dates. We summarise the changes that
make the clones outdated into three types, modified/added/deleted statements
(\textit{S}), file deletion (\textit{D}), and method rewriting (\textit{R}),
along with the issue tracking number and the date of change. The complete set of
101 outdated online clones can be found on the study website.

The outdated online code clones cause problems ranging from uncompilable code
(due to modifications and different API usage in the outdated code) to
the introduction of vulnerabilities to the software~\cite{Xia2014}. An outdated code with a
subtle change (e.g.\ \Cref{fig:before-after}) may be copied and reused without
awareness from developers. Moreover, an outdated code with a defect (e.g.\ a race
condition problem in \Cref{fig:before-after_2}) is clearly harmful to be reused.
Although Stack Overflow has a voting mechanism that may mitigate this issue, the
accepted answer is still used by naive developers who copy and reuse the
outdated code.

@@

To check occurrences of our Stack Overflow outdated code snippets in software projects, 
we

\begin{boxquote}
For RQ4, our results show that 66\% (101) of QS clone pairs on Stack
Overflow are outdated. 86 pairs differ from their newest versions by
modifications applied to variable names or method names, added or deleted
statements, to a fully rewritten code with new method signatures. 15~pairs are
dead snippets.
\end{boxquote}

@@

\begin{table}
\centering
\label{tab:license_github}
\caption{Clones of the 214 missing-license code snippets in 131,703 GitHub projects}
\begin{tabular}{lrr}
	\toprule
	No. of stars & Qualitas & Other Projects \\
	\midrule
	29,540 to 10 stars & 71 & 1,836 \\
	9 to 5 stars & 0 & 739 \\
	4 to 1 stars & 19 & 4,541 \\
	\midrule
	Total & 90 & 7,116 \\
	\bottomrule
\end{tabular}
\end{table}

@@

\begin{boxquote}
For RQ5, we found 214 code snippets on Stack Overflow that
could potentially violate the license of their original software. The majority of them
do not contain licensing statements after they have been copied to
Stack Overflow. For 164 of them, we were able to identify,
with evidence, where the code snippet has been copied from.
\end{boxquote}

@@

On the other hand, the missing of licensing statement of online code clones on
Stack Overflow can cause more damage than outdated code. As shown in our study and
also in the study by An et al. \cite{An2017}, some online clones on Stack Overflow are
originally licensed under more restrictive license than Stack Overflow's CC
BY-SA 3.0. If these missing-license online clones are reused in a software with
incompatible license, the software owner may face legal issues. The software
auditing service such as Black Duck Software or nexB, which can effectively
check for license compliance of code copied from open source projects, cannot
check for the original license of the cloned code snippets on Stack Overflow. Although
Stack Overflow answerers participated in our survey believe that most of the
code snippets on Stack Overflow are too small to claim for copyright and they
fall under fair-use, there is still a risk due to different legal system in each
country. For example, Germany legal system does not have a concept of fair use~\cite{?}
or the number of minimum lines of code to be consider trivial, i.e.~de minimis,
is also different from case to case or from country to country~\cite{?}.

@@

@@

While checking if the copied code snippets on Stack Overflow are still up-to-date with the latest version of their originals can possibly be done automatically, it is a challenging task to automate the process of establishing the origin of code clones. As shown in this study, we still relied on manual validation to read the post, understand the context, and use multiple hints based on human judgement (comments in the code, natural text in the question and answers, date/time of the posts) to conclude that the code snippets are actually copied from Qualitas to Stack Overflow. Thus, one viable solution for now is encouraging the Stack Overflow developers to always include the origin of the copied code snippet in the post so that this link is always established at the posting time. Even better, Stack Overflow can provide an optional form to fill in when an answerer post an answer if he or she copies code from other software projects. The form should include the origin of the code snippet (possibly as a GitHub url) and its original license. Using this manually established links at posting time, we can then automate the process of checking for outdated code.

@@

The implementation and the evaluation of the
above actionable items is part of the agenda of our future work.

>>>>

``Toxic code snippets mean code snippets that, after incorporating into software,
the cost of paying back the technical debt exceeds the value it generates in the long run.'' 
Stack Overflow code snippets
cloned from open source software or online sources can become toxic when they
are (1)~outdated, (2)~violating their original software
license, (3) exhibiting code smells, or (4) having security vulnerabilities. 
We chose to study toxic code snippets caused by outdated code and license-violating code snippets in
this study as the two problems are still under explored compared to code smells~\cite{Tufano2015} or vulnerabilities~\cite{Acar2016,Fischer2017}.
Outdated code snippets can be harmful since they are not up-to-date with their originals and may
contain defects. Code snippets from open source projects usually fall under a
specific software license, e.g.\ GNU General Public License (GPL). If they are
cloned to Stack Overflow answers without the license, and then flow to other projects
with conflicting licenses, legal issues may occur.

@@

\begin{table}
	\centering
	\caption{Stack Overflow and Qualitas datasets}
	\label{tab:datasets}
	%\small
	\begin{tabular}{lrr}
		\toprule
		Data set & No. of files & SLOC \\
		\midrule
		Stack Overflow & 72,365 & 1,840,581 \\ 
		Qualitas &  166,709 & 19,614,083 \\ 
		\bottomrule
	\end{tabular} 
\end{table}

@@

To search for outdated online code clones, we focused on the \textbf{QS} clone
pairs that were cloned from Qualitas to Stack Overflow and compared them with
their latest versions. We downloaded the latest version of the Qualitas projects
from their repositories on October 3, 2017. For each \textbf{QS} online clone
pair, we used the clone from Qualitas as a proxy. We searched for its latest
version by the file name and located the cloned region in the file based on the
method name or specific code statements. We then compared the Stack Overflow
snippet to its latest version line-by-line to find if any change has been made
to the source code. We also made sure that the changes did not come from the
modifications made to the Stack Overflow snippets by the posters but from the
updates in the projects themselves. When we found inconsistent lines between the
two versions, we used {\small\texttt{git blame}} to see who modified those lines
of code and the timestamps. We also read commit messages and investigated the issue tracking information
if the code change is linked to an automatic issue tracking system, such as Jira
or BugZilla to gain insights into the intent behind the change.

Lastly, we searched for the outdated code snippets in 130,703 GitHub projects to
see how widespread is the outdated code in the wild. We mined GitHub
based on the number of stars the projects received, which indicated their popularity. We relied on
GitHub API to query the project metadata before cloning them.
Since GitHub API returned only top 1,000 projects at a time for each query, we formulated the
query to retrieve most starred projects based on their sizes.
The project size range started from 1KB to 2MB with 1KB step size and the last query is
for all the remaining projects that were larger than 2MB.
With this method, we retrieved the top 1,000 most starred projects for each project size. 
As a result, we cloned 130,703 GitHub projects ranging from 29,465 stars to 1 star.
A clone detection was then performed between the outdated code snippets and the GitHub
projects. We selected SourcererCC with 80\% similarity for this task because it
could scale to a large-scale data set, while Simian could not. Finally, we analysed
the clone reports and manually checked the clones.

@@

The visitors survey confirms the findings from the previous studies that Stack
Overflow code snippets can be problematic~\cite{Zhang2018,Acar2016,An2017,Fischer2017}. Sixty-six
percent of the visitors experienced a problem from reusing Stack Overflow code
snippets ranging from incorrect solutions, outdated solutions, mismatched
solutions, to buggy code. Although they are aware of the problems, half of them
(56\%) never reports back to the Stack Overflow discussions. On the other hand,
the visitors rarely give attributions to Stack Overflow when they reuse code
snippets from the website, similar to the findings reported
by~\cite{Baltes2017}. The visitors are generally not aware of the CC BY-SA 3.0
license, and more than half of them never check for license compatibility when
reusing Stack Overflow code snippets. 
Sixty-nine Stack Overflow visitors (79\%) who adopted code from Stack Overflow never 
check if the code snippet originated from a different source (e.g. an open source project) 
and if they had an incompatible license to their projects.
Moreover, we also found that 9\% of the participants
encountered legal issues by copying code from Stack Overflow. To the best of our
knowledge, a study of legal problems from reusing Stack Overflow code snippets
has never been done before and will be our future work.

@@

Moreover, using the information in git commit messages, we can associate each
change to its respective issues in an issue tracking system, such as Bugzilla or
Jira. We found that in 58 cases, the cloned code snippets on Stack Overflow were
changed because of a request in the issue tracking system. Since issue tracking
systems are also used, besides bug reports, for feature request and feature
enhancements, having an issue tracking ID can reflect that most of the changes are important
and not only a superficial fix such as code formatting. The intent behind
the changes are grouped into six categories as shown in \Cref{tab:intent_outdated}.
Enhancement are the majority intent accounting for 65 of the 100 outdated code (65\%), followed 
by code deprecation (15\%). There were 10 outdated code snippets (10\%) due to bug fixing.
The rest of the changes are because of code refactoring (6\%), changing coding style (3\%), and 
data format change (2\%).
Not all outdated code are toxic. However, the 10 buggy and outdated code snippets 
we found are toxic and are harmful for reuse.

\begin{table}
	\centering
	\begin{tabular}{llr}
		\toprule
		Intent & Detail & Amount \\
		\midrule
		Enhancement & Add or update existing features & 64 \\
		Deprecation & Delete dead/deprecated code & 15 \\
		Bug & Fix bugs & 10 \\
		Refactoring & Refactor code for better design & 6 \\
		Coding style & Update to a new style guideline & 3 \\
		Data change & Changes in data format & 2 \\
		\bottomrule
	\end{tabular}
	\label{tab:intent_outdated}
	\caption{Intents of code changes in 100 outdated code snippets}
\end{table}

\Cref{tab:stale_code_details} shows examples of the outdated online clones on
Stack Overflow. The table displays information of the clones from both Stack
Overflow and Qualitas side including the dates. We summarise the changes that
make the clones outdated into three types, modified/added/deleted statements
(\textit{S}), file deletion (\textit{D}), and method rewriting (\textit{R}),
along with the issue tracking number and the date of change. The complete set of
100 outdated online clones can be found on the study website.

@@

\begin{table}[H]
\centering
\label{tab:outdated_github}
\begin{tabular}{lr}
	\toprule
	Clones & Amount \\
	\midrule
	\textit{Found in Qualitas GitHub repos} & 13 \\
	\midrule
	\textit{Found in other project repos} & \\
	Exact copy (outdated) & 47 \\
	Non-exact copy & 32 \\
	\midrule
	Total & 102 \\
	\bottomrule
\end{tabular}
	\caption{Clones of the 100 Stack Overflow outdated code snippets in 131,703 GitHub projects}
\end{table}

Then, we performed the check for clones of the 100 outdated code snippets in 130,703 GitHub projects.
We found 102 cloned candidates appearing in 68 GitHub projects and manually investigated all
of them. As shown in \Cref{tab:outdated_github}, out of the 102 cloned snippets, there were
13 cloned snippets that matched with themselves because some of the Qualitas
projects also appear on GitHub. For other projects besides the Qualitas
projects, there were 47 cloned snippets that were exactly the same as the
outdated code snippets and 42 cloned snippets that were not exactly the same
(e.g.~they contained additional code modifications made by the projects'
developers or they were copied from another source with a slightly different
code).

Among the exact-copy of 47 outdated cloned snippets, there were only two cloned
snippets that gave attributions to Stack Overflow. However, the attributions
pointed to different posts than the ones we found, but containing the same code
in the answers. The other 32 cloned snippets were very likely to be a file-level
clone from its respective original project (e.g. JFreeChart, JUnit, Log4J,
Hadoop) based on their license header and the Javadoc comments. Lastly, the
remaining 13 cloned snippets did not have any hints or evidence of copying.
Twelve of them were toxic because they were the old versions that still contained bugs.

Interestingly, we discovered that the buggy version of the
\texttt{humanReadableInt()} method from Hadoop appears in two popular Java
projects: deeplearning4j (8,830 stars and 4,223 forks) and Apache Hive (1,854
stars and 1,946 forks). Due to the lack of evidence, we could not conclude how
this method, which is exactly the same as the toxic code snippet we found on
Stack Overflow, appears in the two projects. It is possible that the developers
retrieved them from Stack Overflow, other websites, or from Hadoop code base
directly. Nevertheless, we decided to report the developers of the two projects
regarding the issue. We created a bug report for each project (deeplearning4j
\#4694\footnote{deeplearning4j bug report:
	\url{https://github.com/deeplearning4j/deeplearning4j/issues/4692}} and
HIVE-18929\footnote{Apache Hive bug report: \url{
		https://issues.apache.org/jira/browse/HIVE-18929}}) and communicated with the
developers of the projects by describing the problem of race condition in the
outdated version of \texttt{humanReadableInt()} method and proposed a fix by
using the newest version of the method in Hadoop. The issue has been fixed in
deeplearning4j. The developers of deeplearning4j agreed that the method was
problematic and they decided to implement their own fix due to a concern of a
potential software licensing conflict caused by copying the fix directly from
Hadoop code base. On the other hand, the bug report in Apache Hive is still
unfixed.

The outdated online code clones cause problems ranging from uncompilable code
(due to modifications and different API usage in the outdated code) to
the introduction of vulnerabilities to the software~\cite{Xia2014}. An outdated code with a
subtle change (e.g.\ \Cref{fig:before-after}) may be copied and reused without
awareness from developers. Moreover, an outdated code with a defect (e.g.\ a race
condition problem in \Cref{fig:before-after_2}) is clearly harmful to be reused.
Stack Overflow has a voting mechanism that may mitigate this issue, the
accepted answer is still used by naive developers who copy and reuse the
outdated code.
Although we did not find strong evidence of the outdated code snippets
being cloned to GitHub projects, it would still be useful if Stack Overflow
implements a flagging of outdated answers.

\begin{boxquote}
For RQ4, our results show that 66\% (101) of QS clone pairs on Stack
Overflow are outdated. 86 pairs differ from their newest versions by
modifications applied to variable names or method names, added or deleted
statements, to a fully rewritten code with new method signatures. 15~pairs are
dead snippets. 47 outdated code snippets are found in 130,703 GitHub projects
without evidence of copying. Two of them contains a race condition and
are found in popular projects: deeplearning4j and Apache Hive.
\end{boxquote}

@@

@@

The same GitHub study has been done for license-incompatible code snippets. 
We detected clones between the 214
code snippets without their original license (86 QS, 78 EX and 50 UD)  and
130,703 GitHub projects using SourcererCC with 80\% similarity threshold. Then,
we used the Ninka tool to automatically identify software license from the 7,207
code snippets. Opposite to the outdated clones, we discovered many more clones
of 7,207 candidate pairs. There were 90 pairs from 10 Qualitas projects on
GitHub and 7,117 pairs from 2,427 other projects. From
\Cref{tab:license_github}, the clones were found from highly-starred
projects (29,465 to 10 stars) to 1-star projects.
We found 12 code snippets with attributions to Stack Overflow questions/answers and
6 of them were the answers we found as clones from Qualitas or external sources.
Five code snippets did not have license while the originals had Apache-2, GPLv2,
or EPLv1 license. One snippets had AGPLv3 license while the original had Apache-2 license.
Only 995 code snippets in GitHub projects that have the same license
as the originals in Qualitas.

\begin{table}
	\centering
	\label{tab:license_github}
	\begin{tabular}{lrrrrr}
		\toprule
		\multirow{2}{*}{No. of stars} & \multicolumn{2}{c}{Qualitas} & \multicolumn{3}{c}{Other Projects} \\ \cmidrule{2-6}
		& Projects & Pairs & Projects & Pairs & Same license \\
		\midrule
		29,540--10 & 8 & 71 & 406 & 1,837 & 193 \\
		9--5 & 0 & 0 & 275 & 739 & 110 \\
		4--1 & 2 & 24 & 1,746 & 4,536 & 692 \\
		\midrule
		Total & 10 & 95 & 2,427 & 7,112 & 995 \\
		\bottomrule
	\end{tabular}
	\caption{Clones of the 214 Stack Overflow missing-license code snippets in 131,703 GitHub projects}
\end{table}

\begin{boxquote}
For RQ5, we found 214 code snippets on Stack Overflow that
could potentially violate the license of their original software. The majority of them
do not contain licensing statements after they have been copied to
Stack Overflow. For 164 of them, we were able to identify,
with evidence, where the code snippet has been copied from.
We found 7,117 of the 214 license-incompatible code snippets
in 130,703 GitHub projects.
\end{boxquote}

@@

On the other hand, the missing of licensing statement of online code clones on
Stack Overflow can cause more damage than outdated code. As shown in our study and
also in the study by An et al. \cite{An2017}, some online clones on Stack Overflow are
originally licensed under more restrictive license than Stack Overflow's CC
BY-SA 3.0. If these missing-license online clones are reused in a software with
incompatible license, the software owner may face legal issues. The software
auditing service such as Black Duck Software or nexB, which can effectively
check for license compliance of code copied from open source projects, cannot
check for the original license of the cloned code snippets on Stack Overflow. Although
Stack Overflow answerers participated in our survey believe that most of the
code snippets on Stack Overflow are too small to claim for copyright and they
fall under fair-use, there is still a risk due to different legal system in each
country. For example, Germany legal system does not have a concept of fair use
or the number of minimum lines of code to be consider trivial, i.e.~de minimis,
is also different from case to case or from country to country.

@@

\textbf{Automated cloning direction detection:} 
As shown in this study, we still relied on manual
validation to read the post, understand the context, and use multiple hints
based on human judgement (comments in the code, natural text in the question and
answers, date/time of the posts) to conclude that the code snippets are actually
copied from Qualitas to Stack Overflow. We call for more research in
automation of code cloning direction detection. From our experience in this
paper, we found that code comments and accompanied natural text
on Stack Overflow were good source of information to decide the direction of
code copying. Thus, by using code clone detection to locate clone candidates
and then applying information retrieval techniques, e.g. cosine similarity with tf-idf, 
we can rank the clone candidates based on the similarity of their project names (or classes) to the text in 
comments or natural text surrounding the clones in Stack Overflow posts.
For example, the Stack Overflow answer containing the text \textit{``Actually, you can learn how to compare in \textbf{Hadoop}
from \textbf{WritableComparator}. Here is an example that borrows some ideas from it.''}
must be ranked very high among the list of clone candidates of a code snippet
from Hadoop since it contains two terms of the project name (Hadoop) and a class name (WritableComparator)
in it.
This technique will dramatically reduce the manual validation effort to establish
the direction of cloning. The technique can also be used on Stack Overflow to 
flag that an answer has high chance of copying from open source projects.

@@

While checking if the copied code snippets on Stack Overflow are still
up-to-date with the latest version of their originals can possibly be done
automatically, it is a challenging task to automate the process of establishing
the origin of code clones. Since automatically establishing code cloning
is still an open challenge, one viable solution for now is
encouraging the Stack Overflow developers to always include the origin of the
copied code snippet in the post so that this link is always established at the
posting time. Even better, Stack Overflow can provide an optional form to fill
in when an answerer post an answer if he or she copies code from other software
projects. The form should include the origin of the code snippet (possibly as a
GitHub url) and its original license. Using this manually established links at
posting time, we can then automate the process of checking for outdated code.

@@

%The implementation and the evaluation of the
%above actionable items is part of the agenda of our future work.
